---
title: "ClassPrep5"
author: "Jack Fay"
format: 
  html:
    embed-resources: true
---

```{r}
library(tidyverse)
library(here)
theme_set(theme_minimal())
hpi_df <- read_csv(here("Class Prep/data/hpi-tidy.csv"))
hpi_df
```
```{r}
ggplot(data = hpi_df, aes(x = Footprint, y = Wellbeing)) +
  geom_point()
```

```{r}
hpi_us <- hpi_df |> filter(Country == "United States of America")

ggplot(data = hpi_df, aes(x = Footprint, y = Wellbeing)) +
  geom_point() +
  geom_text(data = hpi_us, aes(label = Country)) ## specify
```

```{r}
library(ggrepel)
ggplot(data = hpi_df, aes(x = Footprint, y = Wellbeing)) +
  geom_point() +
  geom_text_repel(data = hpi_us, aes(label = Country)) +
  geom_point(data = hpi_us, size = 3, shape = 1) ## create a second point that is an open circle (shape = 1) with a larger size (size = 3) to surround the United States point on the scatterplot
```

```{r}
ggplot(data = hpi_df, aes(x = Footprint, y = Wellbeing)) +
  geom_point() +
  geom_text_repel(data = hpi_us, aes(label = Country)) +
  geom_point(data = hpi_us, size = 3, shape = 1) +
  labs(title = "Countries with a Higher Ecological Footprint Tend to Have Citizens with Higher Wellbeing", ## add title
    subtitle = "Wellbeing is on a 1-10 scale", ## add subtitle (smaller text size than the title)
    caption = "Data Source: http://happyplanetindex.org/countries", ## add caption to the bottom of the figure
    x = "Ecological Footprint", ## change x axis label
    y = "Wellbeing") ## change y axis label
```


## Exercise 1: For each of the core dplyr functions mutate(), arrange(), select(), slice(), filter(), group_by(), and summarise(), write a one sentence summary of the overall purpose of the function.

mutate() - to create new variables, or edit existing variables in the data set

arrange() - to order cases based on the values of a variable 

select() - to select or remove variables in the data set

slice() - to select certain intervals of cases, specific cases, and to omit certain cases. slice_max() and min allow you to find cases based on variable data

filter() - to select rows based on conditions of the variables

group_by()- to group cases together based on one or more variables

summarise() - creates a new data set where you can examine summary statistics for the existing or transformed data 

## Exercise 2: Review mutate(). Create a new variable in hpi_df that is equal to Wellbeing / Footprint.

```{r}
hpi_df1 <- hpi_df%>%
  mutate(wellbeing_over_footprint = Wellbeing/Footprint)
```


## Exercise 3: Review mutate() and case_when(). Create a new variable in hpi_df that is equal to "80s" if LifeExp is in the 80’s, "70s" if LifeExp is in the 70s, and "Below 70" if LifeExp is less than 70.

```{r}
hpi_df2 <- hpi_df %>%
  mutate(Decade = case_when(
    LifeExpectancy >= 80 ~ "80s",
    LifeExpectancy >= 70 ~ "70s",
    TRUE ~ "Below 70"))
```


## Exercise 4. Review arrange(). Sort the hpi_df data so that the country with the highest LifeExp is in the first row and the country with the lowest LifeExp is in the last row.

```{r}
hpi_df3 <- hpi_df%>%
  arrange(desc(LifeExpectancy))
```


## Exercise 5. Review select(). Make a data frame from hpi_df that only has the variables Country and LifeExp.

```{r}
hpi_df4 <- hpi_df%>%
  select(Country, LifeExpectancy)
```


## Exercise 6. Review slice(). Make a data frame from hpi_df that only has the first 5 rows.

```{r}
hpi_df5 <- hpi_df%>%
  slice(1:5)
```


## Exercise 7. Review filter(). Make a data frame from hpi_df that only has countries in the "Western World" Region.

```{r}
hpi_df6 <- hpi_df%>%
  filter(Region == "Western World")
```


## Exercise 8. Review filter(). Make a data frame from hpi_df that only has countries with a LifeExp less than 55.

```{r}
hpi_df7 <- hpi_df%>%
  filter(LifeExpectancy < 55)
```


## Exercise 9. Review group_by() and summarise(). Make a table of the number of countries in each Region. (Hint: recall that the n() function is the counting function in dplyr).

```{r}
hpi_df8 <- hpi_df%>%
  group_by(Region)%>%
  summarise(total_countries = n())
```


## Exercise 10. Review group_by() and summarise(). Make a table with the maximum LifeExp in each Region.

```{r}
hpi_df9 <- hpi_df%>%
  group_by(Region)%>%
  summarise(highest_life_expectancy = max(LifeExpectancy))
```


Again, we will see non-dplyr functions as well throughout the course, but we will try to review them as they come along. As long as you have a solid dplyr base (and, of course a good ggplot2 base), you should be good to go!